---
title: "Data_Questions"
author: "Ashwin Jayaraman"
date: "11/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```


We haev  6 files for each year. Let us load all the files and combine into 1 file. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
data_2011 <- read_csv('Medicare_Provider_Charge_Inpatient_DRG100_FY2011.csv')
data_2012 <- read_csv('Medicare_Provider_Charge_Inpatient_DRG100_FY2012.csv')
data_2013 <- read_csv('Medicare_Provider_Charge_Inpatient_DRG100_FY2013.csv')
data_2014 <- read_csv("Medicare_Provider_Charge_Inpatient_DRGALL_FY2014.csv")
data_2015 <- read_csv("Medicare_Provider_Charge_Inpatient_DRGALL_FY2015.csv")
data_2016 <- read_csv("Medicare_Provider_Charge_Inpatient_DRGALL_FY2016.csv")
```

Adding the year column to each of the files 

```{r}
data_2011$Year <- 2011
data_2012$Year <- 2012
data_2013$Year <- 2013
data_2014$Year <- 2014
data_2015$Year <- 2015
data_2016$Year <- 2016
```

Now merging the data 


```{r}
data <- rbind(data_2011,data_2012,data_2013,data_2014,data_2015,data_2016)
```

Removing the files loaded from the workspace 

```{r}
rm(data_2011, data_2012, data_2013, data_2014, data_2015, data_2016)
```

Let us look at the data 

```{r}
head(data)
```


Cleaning the amounts in the data 

```{r}
data$`Average Covered Charges` <- gsub('\\$', '', data$`Average Covered Charges`)
data$`Average Covered Charges` <- gsub('\\,', '', data$`Average Covered Charges`)
data$`Average Covered Charges` <- as.numeric(data$`Average Covered Charges`)
data$`Average Medicare Payments` <- gsub('\\$', '', data$`Average Medicare Payments`)
data$`Average Medicare Payments` <- gsub('\\,', '', data$`Average Medicare Payments`)
data$`Average Medicare Payments` <- as.numeric(data$`Average Medicare Payments`)
data$`Average Total Payments` <- gsub('\\$', '', data$`Average Total Payments`)
data$`Average Total Payments` <- gsub('\\,', '', data$`Average Total Payments`)
data$`Average Total Payments` <- as.numeric(data$`Average Total Payments`) 
```

```{r}
data$Total_Payments = data$`Average Total Payments` * data$`Total Discharges`
```


Now some of the key questions that we want to ask is how are payments changing from state to state vis DRG. We have information at a zip code level. So our hypothesis is that certain treatments would be expensive in certain states. 

We also want to see whether certain DRG contribute a lot to the cost. Are there 20% of the DRG which contribute to 80% of the cost. We also want to see the frequencies of these events. Are the costs high for a few number of events which we can term rare events. 

Removing the missing entries 

```{r}
data_na_removed <- na.omit(data)
```

Let us first map the DRG to the relevant Buckets 


```{r}
mapping <- read_csv('Mapping.csv')
data_na_removed$Diagnostic_category <- plyr::mapvalues(data_na_removed$`DRG Definition`, from = mapping$DRG, to = mapping$Name)
```



Let us first identify the percentage of discharges for each of the Diagnostic Categories by Year. So we will be grouping the data by the Diagnostic Categories, the year and the State and identifying the total discharges and the total amount. 

Mutating allows us to give the percentage for each of the diagnostic categories. 

```{r}
drg_grouped_by_year <-  `Inpatient_Cost.(1)` %>% group_by(Diagnostic_category, Year,State) %>% summarise(Count_Discharge = sum(Total.Discharges),Sum_Amount = sum(Total_Payments))

drg_grouped_by_year  <-  drg_grouped_by_year %>% group_by(`Year`) %>% mutate(Percent_Discharges = 100 * Count_Discharge/sum(Count_Discharge),
                                                                             Percent_Cost = 100 * Sum_Amount/sum(Sum_Amount))
```

We have the yearly data by state. We can combine all 6 years by adding to see the total cost for that particular state 

```{r}
drg_grouped_by_year$Average_Amount <- drg_grouped_by_year$Sum_Amount/drg_grouped_by_year$Count_Discharge
```

We also have the average cost by year for each of the states 

Let us now visualize this in the US State Map 

Let us see the Average Cost by the whole state 

```{r}
total_cost_by_state <- drg_grouped_by_year %>% dplyr::group_by(`State`) %>% 
  dplyr::summarise(Total_Discharges = sum(Count_Discharge),Total_Cost = sum(Sum_Amount))

total_cost_by_state$Average_Cost <- total_cost_by_state$Total_Cost/total_cost_by_state$Total_Discharges
colnames(total_cost_by_state)[1] <- 'state'
```

```{r}

library(plotly)
drg_grouped_by_totalAvg <- drg_grouped_by_year %>% group_by(Diagnostic_category) %>%  dplyr::summarise(Total_Discharges = sum(Count_Discharge),Total_Cost = sum(Sum_Amount))


drg_grouped_by_totalAvg$Average_Cost <- drg_grouped_by_totalAvg$Total_Cost/drg_grouped_by_totalAvg$Total_Discharges

g <- ggplot(drg_grouped_by_totalAvg, aes(x = Total_Discharges, y =  Average_Cost)) +
  geom_point(color = "blue") + xlab("Total Discharges") + ylab("Average Cost") + 
  scale_y_continuous(labels = function(l) { paste0("$",l/1000, "K")}) +      scale_x_continuous(labels = function(l) { paste0(l/1000, "K")}) + geom_text(aes(label=Diagnostic_category),hjust=1, vjust=0) +
  theme_dotplot + ggtitle("Average Cost and Discharge Count per DRG")

g

```

```{r}
ggplot(drg_grouped_by_totalAvg, aes(x = Total_Discharges, 
                  y = fct_reorder2(Diagnostic_category, Average_Cost, -Total_Discharges), color = Average_Cost)) + 
    geom_point() + ylab("") + theme_dotplot + scale_x_continuous(labels = function(l) { paste0(l/100000, "K")}) + 
    ggtitle("Number of fatalities per million traffic miles") 
```

```{r}
library(MASS)
drg_grouped_by_totalAvg  <-  drg_grouped_by_totalAvg %>% mutate(Percent_Discharges = 100 * Total_Discharges/sum(Total_Discharges),Percent_Cost = 100 * Total_Cost/sum(Total_Cost))

tidy_form_of_data <- drg_grouped_by_totalAvg %>% gather(key = Category, value = Percentage, -Diagnostic_category) %>% filter(Category %in% c('Percent_Cost', 'Percent_Discharges'))

# ggplot(drg_grouped_by_totalAvg) + geom_point(aes(x = fct_reorder(Diagnostic_category, Percent_Cost), y = Percent_Discharges), color = 'red') + geom_point(aes(x = fct_reorder(Diagnostic_category, Percent_Cost), y = Percent_Cost), color = 'blue') + coord_flip() + theme_dotplot + xlab('Diagnoses Related Groups') + ylab('Percent') + ggtitle('DRGs by Percent Avg Cost & Percent of Discharges') + scale_colour_manual("", breaks = c("Percent_Discharges", "Percent_Cost"), values = c("red", "blue"))


ggplot(tidy_form_of_data, aes(x = Percentage,
                                y = fct_reorder2(Diagnostic_category, fct_reorder(Category,Percentage), -Percentage), color = Category)) +
   geom_point() + ylab("") + theme_dotplot + 
   ggtitle("Number of fatalities per million traffic miles")
```


```{r}
drg_grouped_by_year %>% dplyr::group_by(`State`) %>% 
  dplyr::summarise(Total_Discharges = sum(Count_Discharge),Total_Cost = sum(Sum_Amount))
```


```{r}
library(choroplethr)
library(choroplethrMaps)
```

```{r}
library(Hmisc)
data(state.regions)
head(state.regions)
total_cost_by_state$region <- plyr::mapvalues(total_cost_by_state$state, from = state.regions$abb, to = state.regions$region)
data_na_removed$State <- capitalize(plyr::mapvalues(data_na_removed$`Provider State`, from = state.regions$abb, to = state.regions$region))
```

```{r fig.width = 6}
total_cost_by_state$value <- total_cost_by_state$Average_Cost
state_choropleth(total_cost_by_state,title = "Average Cost by State",
                 legend = "Average Cost")

```

```{r}
drg_grouped_by_year  <-  drg_grouped_by_year %>% group_by(`Year`) %>% mutate(Percent_Discharges = 100 * Count_Discharge/sum(Count_Discharge),         Percent_Cost = 100 * Sum_Amount/sum(Sum_Amount))
```


Alaska has a high verage cost but lets compare that with the total number of discharges and the total cost as well

Let us plot
```{r}

total_cost_by_state$value <- total_cost_by_state$Total_Cost
state_choropleth(total_cost_by_state,title = "Total Cost by State",
                 legend = "Total Cost")

```


```{r}
total_cost_by_state$value<- total_cost_by_state$Total_Discharges
state_choropleth(total_cost_by_state,title = "Total Discharges by State",
                 legend = "Total Discharges")
```

We can see that despite a hugh number of discharges and total vost in the states of AL, MS, Ok. They actually have a very low average cost. On the other hand we see that the states of MT, WY, ND have a low total number of discharges but have a higher average cost. This means that the cost for treating the DRG is pretty expensive in those states. 

Lets us see which DRG is contributing a lot to the cost

```{r}
drg_cost <- drg_grouped_by_year %>% group_by(`Diagnostic_category`) %>% 
  dplyr::summarise(Average_Cost = sum(Sum_Amount)/sum(Count_Discharge))
```


```{r}
#drg_cost$DRG_Definition <- gsub('(.*?)- ', '',drg_cost$`DRG Definition`)

dominant_drg <- drg_cost[order(-drg_cost$Average_Cost),]

top<-head(dominant_drg$Diagnostic_category,10)

top_drg_cost <- drg_cost[drg_cost$Diagnostic_category %in% top,]

```

```{r fig.width = 5}
ggplot(top_drg_cost, aes(x = reorder(Diagnostic_category,Average_Cost) ,y = Average_Cost)) + 
         geom_col() + scale_y_continuous(labels = function(l) { paste0(l/1000, "K")}) +
  coord_flip() + ylab('Average Cost') + xlab('DRG') + ggtitle('Top Cost by DRG')

```

```{r}
drg_discharge <- drg_grouped_by_year %>% group_by(`Diagnostic_category`) %>% 
  dplyr::summarise(Discharge = max(Count_Discharge))

#drg_discharge$DRG_Definition <- gsub('(.*?)- ', '',drg_discharge$`DRG Definition`)

dominant_drg <- drg_discharge[order(-drg_discharge$Discharge),]

top<-head(dominant_drg$Diagnostic_category,10)

top_drg_count <- drg_discharge[drg_discharge$Diagnostic_category %in% top,]

```


```{r}
drg_common <- drg_grouped_by_year %>% group_by(`Diagnostic_category`,`State`) %>% 
  dplyr::summarise(Discharge = sum(Count_Discharge))


most_drg <- drg_common %>%  group_by(`State`) %>%
             filter(Discharge == max(Discharge)) %>%
             arrange(`State`,`Diagnostic_category` )


```


```{r fig.width = 5}
ggplot(top_drg_count, aes(x = reorder(Diagnostic_category,Discharge) ,y = Discharge)) + 
         geom_col() + scale_y_continuous(labels = function(l) { paste0(l/100000, "M")}) +
  coord_flip() + ylab('Count') + xlab('DRG') + ggtitle('Number of Discharges by DRG')
```



Let us now create the cleaveland dot plot to identify the performance of hospitals over the US. We have the state level data and the diagnostic categories. We can create an interactible component which will allow us to do it in a general variable 


```{r}
theme_dotplot <- theme_bw(12) +
    theme(axis.text.y = element_text(size = rel(.75)),
          axis.ticks.y = element_blank(),
          axis.title.x = element_text(size = rel(.75)),
          panel.grid.major.x = element_blank(),
          panel.grid.major.y = element_line(size = 0.5),
          panel.grid.minor.x = element_blank())
```

```{r}

```

```{r}


filter_state <- 'California' 
filter_diagnostic_category <- 'Circulatory System'
state_data <- `Inpatient_Cost.(1)` %>% filter(State == filter_state) %>% filter(Diagnostic_category == filter_diagnostic_category) 

min_num_of_discharges <- 50

provider_information <- state_data %>% group_by(`Provider.Name`) %>% filter(sum(`Total.Discharges`) > min_num_of_discharges) %>%
  summarise(Average_Cost = sum(Total_Payments)/sum(`Total.Discharges`), Discharges = sum(`Total.Discharges`))

number <- 10

costly_providers<-head(provider_information[order(-provider_information$Average_Cost),], number)
cheapest_providers <- head(provider_information[order(provider_information$Average_Cost),], number)

ggplot(costly_providers, aes(x = Average_Cost, y = fct_reorder(`Provider.Name`, Average_Cost))) +
  geom_point(color = "blue") + xlab("Average Cost") + ylab("") + 
  scale_x_continuous(labels = function(l) { paste0("$",l/1000, "K")}) +
  theme_dotplot + 
  ggtitle("Average Cost of Most expensive Providers")


ggplot(cheapest_providers, aes(x = Average_Cost, y = fct_reorder(`Provider.Name`, Average_Cost))) +
  geom_point(color = "blue") + xlab("Average Cost") + ylab("") + 
  scale_x_continuous(labels = function(l) { paste0("$",l/1000, "K")}) +
  theme_dotplot + 
  ggtitle("Average Cost of Least expensive Providers")
  
```




```{r eval=FALSE, include=FALSE}
write.csv(data_na_removed, "Inpatient_Cost.csv")
```

